The seed germination spectrum of alpine plants: a global meta-analysis

### 2020-10-27 doc/manuscript.Rmd
major revision v2.1 ⛰️

---
title: "The seed germination spectrum of alpine plants: a global meta-analysis"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: refs.bib
csl: new-phytologist.csl
link-citations: yes
always_allow_html: true
---

@@

* We present a meta-analysis of the alpine seed germination spectrum based on primary data from germination experiments conducted in four continents with `r MSspecies` plant species, with the aim of testing the generality of alpine germination patterns at the global level.
* We use phylogenetic Generalized Linear Mixed Models to estimate the influence of six environmental cues on seed germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
* Alpine plants are characterized by physiological seed dormancy, strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures. Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also to the lowlands. Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
* Overwintering and warm-cued germination are key drivers of germination in alpine habitats. The interplay of seed morphological traits and germination physiology reflects pressures to avoid frost or drought stress. These results indicate the convergence, at the global level, of the seed germination spectrum of alpine species.

@@

Alpine environments occur worldwide, above the maximum elevation at which trees can grow naturally [@RN3000; @RN4750]. These treeless habitats are characterized by low temperatures, strong winds, unstable substrates and short growing seasons [@RN2392]. Their dependence on extreme and strict climatic conditions [@RN3024], coupled with the relatively small terrestrial area they occupy [@RN4750], makes alpine environments disproportionally vulnerable to climate change [@RN4946; @RN4950]. This threatens the survival of the world’s rich and specialized mountain biodiversity [@RN4947]. It also endangers the sustainability of the ecosystem services that alpine environments provide for hundreds of millions worldwide, including the provisioning of natural resources, the regulation of environmental hazards and the maintenance of cultural and recreational activities [@RN4949; @RN4948]. Succesful management of climate change and of its consequences will depend on an adequate understanding of alpine biodiversity.

Specially, plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for active efforts in plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241]. Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by seed is the main strategy to maintain genetic diversity and colonize suitable new sites in response to environmental changes. However, the short reproductive season of the alpine habitat constrains the phenological timing of flowering, pollination, seed development and seed dispersal. For regeneration by seed to be succesful, dispersed propagules must germinate in the most appropriate period to ensure the survival of seedlings under environmental stress [@RN4714; @RN2868; @RN4717]. The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691]. In alpine habitats, seed germination will be a crucial test for plant populations coping with climate change, determining whether they will persist or go into extinction.

Natural selection is expected to favour seed germination requirements that reduce the probability of facing environmental conditions which are not appropriate for seedling survival and growth [@RN3375]. Thus, germination is mostly regulated by environmental cues related to water availability and soil temperature [@RN3368]. The propagules of many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214]. Different degrees of dormancy within a seed population work to ensure the distribution of seedling emergence across time, thereby bet hedging offspring survival against unpredictable environments [@RN3065]. Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719]. Given the comparatively high heterogeneity of alpine climates and plant lineages adapted to different regions, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants while accounting for the evolutionary relatedness among taxa. The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats. A major goal for trait-based seed ecology is to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].

In the available literature, many alpine species have been described as having seeds with deep physiological dormancy, i.e. a type of dormancy that is caused by the seed’s internal balance of phytohormones and requires a long exposure to dormancy-breaking treatments to be terminated [@RN2943; @RN3703; @RN3214]. This physiological dormancy would prevent precocious germination of recently dispersed seeds during autumn, when appropriate conditions are not likely to persist for more than a few weeks or days [@RN3330]. A requirement for cold stratification to break dormancy would allow seeds to sense the snow season, thereby postponing germination to a more favourable period for seedling survival and development. Thus, it can be expected that a cold stratification period is a common requirement for seed germination in alpine plants. Nonetheless, an early review [@RN3213] suggested that cold stratification was not a requirement for the seed germination of several alpine species from different mountain ranges in the USA, a finding that has been reported by subsequent studies [@RN1484; @RN2995; @RN4721]. More recently, research on Australian alpine plants found that a cold stratification period only increased germination in half of the species tested [@RN3703; @RN3285]. In contrast, cold stratification increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707]. Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate even after cold stratification [@RN4713]. Finally, in a comparative study on germination of unstratified seeds from different habitats of the Austrian Alps, species from alpine grasslands had the lowest final germination proportion, suggesting that they had the highest level of dormancy [@RN4930].

Seed dormancy is not the only trait that can prevent, in practice, autumn germination. Early studies on the germination ecology of alpine plants demonstrated that in most species recently dispersed seeds require relatively high temperatures for germination [@RN3258; @RN3213; @RN4712]. Warm-cued germination has been considered as an adaptation to prevent seed germination at the time of seed dispersal (autumn) when temperatures are relatively low and there is a high risk of frost [@RN3393]. Indeed, germination of alpine seeds tends to occur after winter, mainly in early summer [@RN2392; @RN2382]. In recent times, an increasing number of studies reported that fresh seeds of some alpine plants also germinate at cool incubation temperatures or during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713]. This cold-cued germination could be an adaptation to germinate under snow or during snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], or to attain an optimal size for overwintering [@RN4712]. Yet, the generality of warm-cued germination across different alpine regions and habitats has never been studied systematically, and the relative importance of cold stratification versus warm-cued germination remains unclear. A central question is whether physiological seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or the warm-cued germination is enough to prevent germination under cool autumn temperatures. The response to this question becomes pertinent under a warming climate, which could result in non-dormant seeds germinating readily in autumn. Non-dormant seeds could possibly have an advantage as the growing seasons become longer, if seedlings have enough time to grow and become sufficiently large to survive the winter snow season. Potentially, having a non-dormant or less dormant seed could be a better strategy for alpine plants to thrive under future climate [@RN3317; @RN2382]. However, we might also expect that seedlings from these species will be exposed to frost damage if snow starts to melt earlier and insulation during winter disappears [@RN4944] and therefore a mechanism that enables diverse germination timing (i.e. low germination synchrony) should be preferred.

Besides patterns of seed dormancy and germination temperature, the response of alpine seeds to diurnal alternating temperatures and light can provide further information on their capacity to detect fine-scale environmental cues. Fluctuations of diurnal temperature decrease with increasing burial depths; below 10 cm soil temperature might be too stable to trigger germination in species requiring temperature alternation [@RN4910]. Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909]. Therefore, these two cues can indicate whether seeds are buried deep in the soil, in rock crevices or under snow. The depth of seed burial is crucial for seedling emergence, as small seeds lack the resources to survive until they reach the soil surface [@RN4898]. A study of 445 species from the Qinghai-Tibet plateau found that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376]. This is intriguing given that alpine habitats typically have a strong day-night climatic variation during the growing season. Regarding light and darkness, some local studies have found alpine species to have a preference for germinating in light [@RN4730; @RN4729]. This trait could favour the creation of a persistent soil seed bank [@RN4718; @RN4929] which is advantageous in the temporally and spatially unpredictable alpine environments [@RN4731; @RN4732]. Indeed, alpine soil seed banks are more frequent than previously supposed, as consistently reported in recent years [@RN4929; @RN4718; @RN4945]. Nevertheless, other alpine species seem to prefer dark germination [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008]. Although some germination studies report laboratory experiments conducted with constant and alternating temperatures, or light and dark conditions, they are very much case-dependent, and therefore it is difficult to evaluate whether these factors show a consistent pattern across global alpine ecosystems.

In this work, we make a global collection of primary germination data from laboratory experiments conducted with seeds of alpine plant species. The assembled dataset represents eight alpine regions belonging to four continents, and we use it to investigate the germination response of alpine plants to key environmental factors. These factors include dormancy-breaking treatments (cold stratification, GA~3~ and scarification), average germination temperature, diurnal alternating temperatures and light [@RN3368; @RN3214]. Given the heterogeneity of alpine species lineages adapted to different regions, we account for the shared evolutionary history using a phylogenetic comparative approach [@RN3081]. We also analyse the relationship between germination ecology and two morphological traits which are key drivers of the seed ecological spectrum: seed mass and the embryo to seed ratio [@RN4655]. Our main aim in conducting this work is to test the consistency of general assumptions about the influence of major environmental cues on seed germination of alpine plants. Specifically, we determine at the global scale whether alpine seeds: (1) have physiological dormancy; (2) require cold stratification to break dormancy; (3) need relatively warm temperatures to germinate; (4) do not require alternating temperatures for germination; and (5) respond positively to light. To understand the ecological constraints of our meta-analysis on alpine species pools, we analyse how the germination responses differ between strict alpine species (that occur exclusively above the treeline) and generalist species (that occur above the treeline but also in the lowlands). We can expect that the former group will show a more convergent alpine germination syndrome, while the latter will be more plastic in their response.

@@

We conducted a meta-analysis of primary data [@RN4734], which consists of pooling together original data from different local studies to synthesize their conclusions at the global level. All data processing, analysis and manuscript production has been performed in *R* [@RN2315]. The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds. (The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review). To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments. The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated. We received data (Table 1) from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Northern Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps. We also received data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).

@@

Once received, we standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R*. For each species, we recorded its elevation range from local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*, https://www.anbg.gov.au/cgi-bin/anhsir). We used the elevation ranges (i) to remove from the dataset outlier species that do not grow above the treeline; and (ii) to classify the rest of the species as *strict alpine* (i.e. species that only grow above the treeline) or *generalist* (i.e. species that can grow above the treeline but also in the lowlands). For each species, we compiled three additional traits. First, seed dormancy class obtained from Baskin & Baskin [-@RN3214] and Rosbakh et al. [-@RN4743]. We simplified the original dormancy classes to five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant. Second, seed mass as provided by the Royal Botanic Gardens Kew Seed Information Database [-@RN2987] and some additional sources [@RN4728; @RN4735; @RN3376]. Third, the embryo to seed surface ratio. Embryo to surface area values were based on measurements of the surface area of the embryo and seed from drawings and photographs retrieved mainly from Martin [-@RN3244] and the Royal Botanic Gardens Kew Seed Information Database [-@RN2987], supplemented with own measurements (Vandelook, unpublished). For some species that were missing seed mass or embryo values, we calculated genus or family averages.

@@

We used the final germination proportion (i.e. the number of seeds germinated out of seeds sown) at 8 weeks as the main response variable for the germination treatments. We further calculated the mean germination time and the germination uncertainty index with the package *GerminaR* [@RN4752]. Mean germination time is the time that it takes for half of the seed lot to germinate, indicating the germination speed in each experimental condition. The germination uncertainty index estimates how scattered germination is through time, and gives an idea of whether germination is synchronous (i.e. most seeds germinate around the mean germination time) or asynchronous (i.e. germination events are separated through time during the germination experiment). Lower values in the germination uncertainty index indicate higher germination synchrony. The uncertainty index is analogous to the synchrony index and we used it because, unlike the latter, its values are not constrained between 0 and 1 [@RN4752]. To calculate these indices, we used only records that met two conditions: (i) more than 50% final germination percentage; and (ii) a mean germination time under 4 weeks. We established these limits to prevent the results being altered by a modification of the seed’s dormancy status during the germination incubation (e.g. a fraction of the seed lot is non-dormant at the beginning of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).

@@

We meta-analysed germination data by fitting generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755]. To model final germination proportion, we used binomial MCMCglmms, while for mean germination time and the germination uncertainty index we used gaussian MCMCglmms. Models had, as fixed effects, the experimental conditions (stratification, GA~3~, scarification, temperature, alternating temperature and light), plus their interaction with seed mass and embryo:seed, and the strict alpine or generalist character of the species. Random effects included a reconstructed phylogenetic tree for the `r MSspecies` species, and species identity, seed lot, lab (i.e. data provider), and alpine region (Table 1). To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754]. We placed taxa absent from the mega-tree at the genus-level basal node. In all models, all variables were scaled so their contribution to the effect sizes could be compared. We used weakly informative priors in all models, with parameter-expanded priors for the random effects. Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions. From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI). We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant. To estimate the phylogenetic signal of seed germination over all variables we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al. [-@RN4756]. When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757].

@@

The final dataset used in our meta-analysis contained `r MSrecords` records (i.e. germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species (`r MSstrict` strict alpine species and `r MSgeneral` generalist species). Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), in both hemispheres.

Most species in the dataset were reported as having dormant seeds (Fig. \@ref(fig:fig1)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species. Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morpho-physiological and then physical. Both strict alpine and generalist species had a similar range of values for seed mass (Fig. \@ref(fig:fig1)b) and embryo:seed ratio (Fig. \@ref(fig:fig1)c). The values of seed mass ranged from `r MSminMass` mg (*`r MSminMassSpp`*) to `r MSmaxMass` mg (*`r MSmaxMassSpp`*), with a median value of `r MSmedianMass` mg. Embryo:seed ratio encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds (`r MSmaxEmbryo`). The median value was `r MSmedianEmbryo`.

>>>>

---
title: "The seed germination spectrum of alpine plants: a global meta-analysis"
output: 
  bookdown::word_document2:
    number_sections: false
    reference_docx: template.docx
bibliography: refs.bib
csl: new-phytologist.csl
link-citations: yes
always_allow_html: true
---

@@

* Given the micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology, a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
* We meta-analyse primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on seed germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
* Alpine plants show physiological seed dormancy, a strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures. Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also to the lowlands. Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
* Globally, overwintering and warm-cued germination are key drivers of germination in alpine habitats. The interplay between germination physiology and seed morphological traits further reflects pressures to avoid frost or drought stress. Our results indicate the convergence, at the global level, of the seed germination spectrum of alpine species.

@@

Alpine environments occur worldwide, above the maximum elevation at which trees can grow naturally [@RN3000; @RN3024; @RN4750]. These treeless habitats are characterized by low temperatures, unstable substrates and short growing seasons [@RN2392]. Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by diaspores (i.e. seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and colonize suitable new sites in response to environmental changes. Plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241]. However, the short reproductive season of the alpine habitat constrains the phenological timing of flowering, pollination, seed development and seed dispersal. For regeneration by seed to be successful, dispersed propagules must germinate in the most appropriate period to ensure the survival of seedlings [@RN4714; @RN2868; @RN4717]. The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691]. In alpine habitats, seed germination will be a crucial test for plant populations coping with climate change, determining whether they will persist or go into extinction.

Natural selection is expected to favour seed germination requirements that reduce the probability of facing environmental conditions which are not appropriate for seedling survival and growth [@RN3375]. Thus, germination is mostly regulated by environmental cues related to water availability and soil temperature [@RN3368]. The propagules of many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214]. Different degrees of dormancy within a seed population ensure the distribution of seedling emergence across time, a form of bet-hedging offspring survival against unpredictable environments [@RN3065]. Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719]. Given the high micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants. The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats. A major goal for trait-based seed ecology is therefore to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].

Traditionally, alpine species have been described as having seeds with physiological dormancy, i.e. a type of dormancy that is caused by the seed’s internal balance of phytohormones and requires a long exposure to dormancy-breaking treatments to be terminated [@RN2943; @RN3703; @RN3214]. This physiological dormancy would prevent precocious germination of recently dispersed seeds during autumn, when appropriate conditions are not likely to persist for more than a few weeks or days [@RN3330]. A requirement for cold stratification to break dormancy allows seeds to sense the snow season, thereby postponing germination to a more favourable period for seedling survival and development. Thus, it can be expected that a cold stratification period is a requirement for seed germination in alpine plants. One of the first experiments on this topic was conducted by Söyrinki [-@RN4951], who showed how the storage of arctic-alpine species at winter temperatures (i.e. cold stratification) increased seed germination in most species. Nonetheless, Amen [-@RN3213] later suggested that cold stratification was not a requirement for the seed germination of alpine species from different mountain ranges in the USA, a finding that was supported by subsequent studies [@RN1484; @RN2995; @RN4721]. More recently, research on Australian alpine plants found that a cold stratification period only increased germination in half of the species tested [@RN3703; @RN3285]. Cold stratification also increased seed germination over a range of temperatures in most of the species studied in the alpine zone of Japan [@RN707]. Along an elevational gradient in the central Chilean Andes, cold stratification was an important requirement for seed germination in species from lower elevations, while species from higher sites did not germinate after a single period of cold stratification [@RN4713]. Finally, in a comparative study on germination of unstratified seeds from different habitats of the Austrian Alps, species from alpine grasslands had the lowest final germination proportion, suggesting that they had the highest level of dormancy [@RN4930].

Seed dormancy is not the only trait that can prevent, in practice, autumn germination. Early studies on the germination ecology of alpine plants demonstrated that in most species recently dispersed seeds require relatively high temperatures for germination [@RN4951; @RN3258; @RN3213; @RN4712]. Warm-cued germination has been considered as an adaptation to prevent seed germination at the time of seed dispersal (autumn) when temperatures are relatively low and there is a high risk of frost [@RN3393]. Indeed, germination of alpine seeds tends to occur after winter, mainly in early summer [@RN2392; @RN2382]. In recent times, an increasing number of studies reported that fresh seeds of some alpine plants can also germinate at cool incubation temperatures or during cold stratification treatments [@RN2943; @RN3285; @RN2371; @RN4713]. This cold-cued germination could be an adaptation to germinate under snow or during snowmelt, which could presumably allow seedlings to develop a deeper root system before topsoil desiccation in summer [@RN4720], or to attain an optimal size for overwintering [@RN4712]. Yet, the generality of warm-cued germination across different alpine regions and habitats has never been studied systematically, and the relative importance of cold stratification versus warm-cued germination remains unclear. A central question is whether physiological seed dormancy plays a relevant role in postponing alpine germination during autumn and winter, or whether the warm-cued germination is enough to prevent germination under cool autumn temperatures. The response to this question becomes pertinent under a warming climate, which could result in non-dormant seeds germinating readily in autumn. Non-dormant seeds could possibly have an advantage as the growing seasons become longer, if seedlings have enough time to grow and become sufficiently large to survive the winter snow season. Potentially, having a non-dormant or less dormant seed could be a better strategy for alpine plants to thrive under future climate [@RN3317; @RN2382]. However, we might also expect that seedlings from these species will be exposed to frost damage if snow starts to melt earlier and insulation during winter disappears [@RN4944] and therefore mechanisms that enable diverse germination timing (i.e. low germination synchrony) may be favoured.

Besides patterns of seed dormancy and germination temperature, the response of alpine seeds to diurnal alternating temperatures and light can provide further information on their capacity to detect fine-scale environmental cues. Fluctuations of diurnal temperature decrease with increasing burial depths and the depth profile of this decrease depends on soil type, vegetation cover and snow [@RN1380; @RN4910; @RN4952]. Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909]. Therefore, the response to alternating temperatures and light can indicate seeds whether they are buried deep in the soil, in rock crevices or under snow. The depth of seed burial is crucial for seedling emergence, as small seeds lack the resources to survive until they reach the soil surface [@RN4898]. A study of 445 species from the Qinghai-Tibet plateau found that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376]. This is intriguing given that alpine habitats typically have a strong day-night climatic variation during the growing season. Regarding light and darkness, some local studies have found alpine species to have a preference for germinating in light [@RN4730; @RN4729]. This trait could favour the creation of a persistent soil seed bank [@RN4718; @RN4929] which is advantageous in the temporally and spatially unpredictable alpine environments [@RN4731; @RN4732]. Indeed, alpine soil seed banks are more frequent than previously supposed, as consistently reported in recent years [@RN4929; @RN4718; @RN4945]. Nevertheless, other alpine species seem to prefer dark germination [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008]. Although some germination studies report laboratory experiments conducted with constant and alternating temperatures, or light and dark conditions, they are very much case-dependent, and therefore it is difficult to evaluate whether these factors show a consistent pattern across global alpine ecosystems.

In this work, we make a global collection of primary germination data from laboratory experiments conducted with seeds of alpine plant species from extratropical regions. The assembled dataset represents eight seasonal alpine regions belonging to four continents. We meta-analyse this dataset to investigate the germination response of alpine plants to key environmental factors: dormancy-breaking treatments (cold stratification, GA~3~ and scarification), average germination temperature, diurnal alternating temperatures and light [@RN3368; @RN3214]. Given the heterogeneity of alpine species lineages adapted to different regions, we account for the shared evolutionary history using a phylogenetic comparative approach [@RN3081]. We further analyse the relationships between germination ecology and two morphological traits which are key drivers of the seed ecological spectrum: seed mass and the embryo to seed ratio [@RN4655]. Our main aim in conducting this work is to test the consistency of general assumptions about the influence of major environmental cues on seed germination of alpine plants. Specifically, we determine at the global scale whether alpine seeds: (1) have physiological dormancy; (2) require cold stratification to break dormancy; (3) need relatively warm temperatures to germinate; (4) do not require alternating temperatures for germination; and (5) respond positively to light. To understand the ecological constraints of our meta-analysis on alpine species pools, we compare the germination responses of strict alpine species (that occur exclusively above the treeline) and generalist species (that occur above the treeline but also at lower elevations). We can expect that the former group will show a more convergent alpine germination syndrome, while the latter will be more plastic in their response.

@@

We conducted a meta-analysis of primary data [@RN4734], which consists of pooling together original data from different local studies to synthesize their conclusions at the global level. All data processing, analysis and manuscript production has been performed in *R* [@RN2315]. The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds. (The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review). To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments. The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated. We received data (Table 1) on species from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Northern Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps. We also received data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).

@@

We standardized the species names according to *The Plant List* [@RN2321] using the package *Taxonstand* [@RN4660] in *R*. For each species, we recorded its elevation range from local floras [@RN4741; @RN4739; @RN3243; @RN4742; @RN4740], catalogues [@RN4738] and herbaria (*Australian National Herbarium*, https://www.anbg.gov.au/cgi-bin/anhsir). We used the elevation ranges (i) to remove from the dataset outlier species that do not grow above the treeline; and (ii) to classify the rest of the species as *strict alpine* (i.e. species that only grow above the treeline) or *generalist* (i.e. species that can grow above the treeline but also in the lowlands). For each species, we compiled three additional traits: seed dormancy class, seed mass and embryo:seed ratio.

We obtained seed dormancy class from Baskin & Baskin [-@RN3214] and Rosbakh et al. [-@RN4743]. We simplified the original dormancy classes to five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant. We followed the concept by Vleeshouwers et al. [-@RN3254] to define dormancy as an inner seed property (sometimes called "endodormancy"), which should be separated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al. [-@RN2877] which considers morphological and physical dormancy as classes separated from physiological dormancy.

We collected mean seed mass for each species as provided by the Royal Botanic Gardens Kew Seed Information Database [-@RN2987] and additional sources [@RN4728; @RN4735; @RN3376]. We recorded the embryo to seed surface area ratio for each species. Embryo to seed surface area values were calculated by dividing embryo area by the surface area of the seed, more specifically embryo plus endosperm and perisperm. Measurements were made using the *ImageJ* software [@RN4708] on drawings and photographs of seeds cut in half along the longitudinal axis, retrieved mainly from Martin [-@RN3244] and the Royal Botanic Gardens Kew Seed Information Database [-@RN2987], supplemented with own photographs (Vandelook, unpublished). For some species that were missing seed mass or embryo values, we calculated genus or family averages.

@@

We used the final germination proportion (i.e. the number of seeds germinated out of seeds sown) at 8 weeks as the main response variable for the germination treatments. Eight weeks was a compromise deadline to unify datasets, as there was a small fraction of trials that had run for longer times. We further calculated the mean germination time and the germination uncertainty index with the package *GerminaR* [@RN4752]. Equations for these two germination indices are available in Lozano-Isla et al. [-@RN4752]. Mean germination time is the time that it takes for half of the seed lot to germinate, indicating the germination speed in each experimental condition. The germination uncertainty index estimates how scattered germination is through time, giving an idea of whether germination is synchronous (i.e. most seeds germinate around the mean germination time) or asynchronous (i.e. germination events are separated through time during the germination experiment). Lower values in the germination uncertainty index indicate higher germination synchrony. The uncertainty index is analogous to the synchrony index and we used it because, unlike the latter, its values are not constrained between 0 and 1 [@RN4752]. To calculate these indices, we used only records that met two conditions: (i) more than 50% final germination percentage; and (ii) a mean germination time under 4 weeks. Within these limits, we were able to calculate mean germination time and uncertainty for `r MSfamdspp` species. We established this conservative approach to prevent the results being altered by a modification of the seed’s dormancy status during the germination incubation (e.g. a fraction of the seed lot is non-dormant at the beginning of the incubation and germinates shortly after the start of the experiment; while another fraction is originally dormant, loses dormancy in response to the incubation conditions, and germinates in a second wave).

@@

We meta-analysed germination data by fitting generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755]. To model final germination proportion, we used binomial MCMCglmms, while for mean germination time and the germination uncertainty index we used gaussian MCMCglmms. Models had, as fixed effects, the experimental conditions (stratification, GA~3~, scarification, temperature, alternating temperature and light), plus their interaction with seed mass and embryo:seed, and the strict alpine or generalist character of the species. Random effects included a reconstructed phylogenetic tree for the `r MSspecies` species, and species identity, seed lot, lab (i.e. data provider), and alpine region (Table 1). To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754]. We placed taxa absent from the mega-tree at the genus-level basal node. In all models, all variables were scaled so their contribution to the effect sizes could be compared. We used weakly informative priors in all models, with parameter-expanded priors for the random effects. Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions. From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI). We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant. To estimate the phylogenetic signal of seed germination over all variables, we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al. [-@RN4756]. When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757].

@@

The final dataset used in our meta-analysis contained `r MSrecords` records (i.e. germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species. From these, `r MSstrict` were categorized as strict alpine species and `r MSgeneral` as generalist. Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), in both hemispheres. All regions correspond to middle latitudes.

Most species in the dataset were reported as having dormant seeds (Fig. \@ref(fig:fig1)a), but the frequency of the dormancy classes did not differ between generalist and strict alpine species. Among those species with dormant seeds, physiological dormancy was the most common dormancy class, followed by morphophysiological and then physical. Both strict alpine and generalist species had a similar range of values for seed mass (Fig. \@ref(fig:fig1)b) and embryo:seed ratio (Fig. \@ref(fig:fig1)c). The values of seed mass ranged from `r MSminMass` mg (*`r MSminMassSpp`*) to `r MSmaxMass` mg (*`r MSmaxMassSpp`*), with a median value of `r MSmedianMass` mg. Embryo:seed ratio encompassed the full range of potential values, from endospermic species with very small embryos (`r MSminEmbryo`) to non-endospermic seeds that store nutrients in the cotyledons (`r MSmaxEmbryo`). The median value was `r MSmedianEmbryo`.
